Qualifying Report

### 2019-01-08 Source/QualifyingReport/QualifyingReport.tex
work done overnight on QR

% HEADER BEGINS
\documentclass{article}
%% Packages
\input{settings/packages}
\usepackage{graphicx}
\usepackage{float}
\usepackage{parskip}
%% Page Settings
\input{settings/page}
%% Own Commands
\input{settings/macros}
\title{Qualifying Report}
\date{2018-12-30}
\author{Jomar Alcantara}
% HEADER ENDS
% DOCUMENT BEGINS
\begin{document}
\maketitle

@@

@@

@@

\section{Project Proposal}
\textbf{Full Provisional Title of the Research:} Title here \newline
\textbf{Research Questions and Hypotheses}
The broad question is can we use Machine Learning to assist in the early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment. There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia. These studies have identified a number of speech and language differences between healthy controls and those with dementia. Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia. However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'. While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments. \newline
\par
Machine Learning itself is undergoing a shift in focus at the present time. I will describe this shift using the context of my proposed research. In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density. These features are then used to classify patients. A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score. \newline
\par
Deep Learning takes a different approach by bypassing the feature generating stage. Deep learning uses neural networks as a fundamental building block.

\subsection{Aims and Rationale}
The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests. There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia. Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other. I'm not going to control for memory problems as a potential confound, but does this weaken the research? How do I defend this? \newline
\begin{itemize}
	\item Hypothesis. I predict that, given a number of samples of language over a given time from, is it possible to detect language decline in humans such that you can differentiate between normal language decline and decline in those with MCI.
	\item Hypothesis. I predict that my new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
	\item My first hypothesis is whether monitoring language samples and other measures of cognitive decline over the course of the year is enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
	\item My second hypothesis is that those with MCI show a different language profile to controls, if matched similarly for education and other significant factors.
\end{itemize}	
\section{Experiment 1 - Pilot Study of Neuropsychological Tests}
\subsection{Methods}

@@

\subsection{Results}
\subsection{Discussion}
\subsection{Conclusion}
\section{Experiment 2 - Initial Study of the Presidents corpus}
\subsection{Background}
\subsection{Methods}

@@

\subsection{Results}
\subsection{Discussion}
\subsection{Conclusion}

\subsection{Timeline of Future Work and Publication plan}
\textbf{Year One} \newline
\textbf{Year Two} \newline
\textbf{Year Three} \newline

\subsection{Provisional Table of Contents}
\subsubsection{Chapter 1: Introduction}
Introduction to the problem of dementia in the context of the wider world including quality of life and financial implications. Exploration of dementia as a syndrome rather than a disease, and a look at the different variants of dementia. A look at the rationale behind research into the early diagnosis of dementia as well as a brief look at what has been done in the area (wide context, so pharmacological and psychological). \newline
\subsubsection{Chapter 2: Background}
Literature Review which looks at the background of what has been done in the area of early diagnosis of dementia, from a purely psychological perspective as well as the work done in machine learning in this topic. Look at the different attributes that researchers have explored. Look at the difference between traditional machine learning methods and deep learning and discuss the pros and cons of each perspective. \newline
\subsection{Chapter 3: Experiments with existing datasets: Presidential Press Conferences, Three Authors - Murdoch, Christie and James, DementiaBank dataset.}
Aim: To replicate and extend the work by Berisha and Liss. To use traditional machine learning techniques to test whether they can differentiate between Reagan, Bush and Trump using similar features. To test to see whether Deep Learning can also differentiate between the three. \newline
Dataset: the presidential press conferences given by Presidents Ronald Reagan, George H.W. Bush and Donald Trump. \newline
Paper: This experiment replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative comparison to Ronald Reagan. This experiment will look at the features originally recommended by Berisha and Liss, as well as any others that have potential as discussed in the literature review above. \newline
Aim: To replicate the work by Le et al. To use traditional machine learning techniques to see if it can differentiate between the three authors using similar features. To test to see whether deep learning can also differentiate between the three. To add 'pre-trained' layers from Presidential Debates experiment to see if this improves accuracy. \newline
Dataset is the literary novels of Iris Murdoch, Agatha Christie and P.D. James. \newline
Paper:
Aim: To replicate and extend the work of Orimaye et al. To use traditional machine learning techniques to see if it can categorise people into dementia or healthy categories using the samples given. To see whether can use pre-trained layer from presidential debates to improve accuracy. It doesn't seem relevant to use pretrained layer from Experiment four, but it might be worth looking at the results of this. \newline
Dataset: The DementiaBank Corpus. \newline
Paper: To replicate and extend the work of Orimaye et al. To compare traditional machine learning techniques (already completed by Orimaye) and deep learning. \newline
\subsection{Chapter 4: Validation of new repeatable neuropsychological battery for detecting MCI.}
Aim: \newline
Dataset: To be created.

\subsection{Chapter 5: Longditunal Study }
\subsubsection{Dataset creation}
The aim is to recruit a set of 50 participants with MCI / Early Dementia (diagnosed or undiagnosed) and a set of 50 controls with a similar age and broadly similar educational background (a known potential confound in assessing cognitive decline). The rationale for recruiting this number is that there will be an expected drop out rate and so
Paper:
\subsubsection{Ethics}
Need to discuss confidentiality right at the start of each conversation and to reiterate at every meeting. I will also provide an information sheet which details the confidentiality statement. The spoken confidentiality statement will be as follows.
"Anything you say to me today will be confidential unless I feel that there is a risk to you, or someone else. If this is the case, I may need to speak to or write to third parties such as your GP. Is this ok?" \newline
\par
\subsubsection{What if someone reveals suicidal ideation or other risks?}
As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9). This is something to explore further. As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.  Optionally with the tape recorder turned on, or off as requested.  Any text recorded during the risk assessment will not be included in the language analysis. Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided. A discussion will then be made about whether it is appropriate to notify the GP of the disclosure. In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact. (Crisis Team and GP). In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken. \newline
\subsubsection{What happens at the end of the data collection phase if there is a suspected decline in a participant's cognitive function}
\subsection{How does language differ between those with AD, MCI and Controls}
\subsection{Chapter 7: Overall results and Discussion}
\subsection{Chapter 8: Conclusions and Future work}

>>>>

% HEADER BEGINS
\documentclass{article}
%% Packages
\input{settings/packages}
\usepackage{graphicx}
\usepackage{float}
%%\usepackage{parskip}
\usepackage{enumitem}
\usepackage{fancyhdr}
%% Page Settings
\input{settings/page}
%% Own Commands
\pagestyle{fancy}
\fancyhf{}
\fancyhead[CE,CO]{Early Detection of MCI - An Interdisciplinary Approach}
\fancyfoot[CE,CO]{\leftmark}
\fancyfoot[LE,RO]{\thepage}

\renewcommand{\headrulewidth}{2pt}
\renewcommand{\footrulewidth}{1pt}
\input{settings/macros}
\title{Qualifying Report}
\date{2019-01-07}
\author{Jomar Alcantara}
% HEADER ENDS
% DOCUMENT BEGINS
\begin{document}
\maketitle
\newpage
\begin{abstract}

\end{abstract}

@@

\listoftables
\listoffigures

@@

\subsection{What is this subsection called?}
\textbf{Full Provisional Title of the Research:} Title here \newline
\textbf{Research Questions and Hypotheses}
The broad question is can we use Machine Learning to assist in the early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment. There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia. These studies have identified a number of speech and language differences between healthy controls and those with dementia. Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia. However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'. While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments. \newline
Machine Learning itself is undergoing a shift in focus at the present time. I will describe this shift using the context of my proposed research. In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density. These features are then used to classify patients. A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score. \newline
Deep Learning takes a different approach by bypassing the feature generating stage. Deep learning uses neural networks as a fundamental building block.

\subsection{Aims and Rationale}
The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to to time-consuming and expensive psychological tests. There is a lot of research into the analysis of language as a bio-marker for MCI and Early Dementia. Given that sampling a person's language is relatively effortless, my research looks at whether we can find bio-markers of MCI and Early Dementia in natural language.\newline
Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other. I'm not going to control for memory problems as a potential confound, but does this weaken the research? How do I defend this? \newline
\begin{itemize}
	\item Hypothesis. I predict that, given a number of samples of language over a given time from, is it possible to detect language decline in humans such that you can differentiate between normal language decline and decline in those with MCI.
	\item Hypothesis. I predict that my new neuropsychological battery of tests will be more sensitive at detecting mild cognitive impairment than existing measures.
	\item My first hypothesis is whether monitoring language samples and other measures of cognitive decline over the course of the year is enough to detect cognitive decline in a population of those with Subjective Cognitive Decline and Mild Cognitive Impairment.
	\item My second hypothesis is that those with MCI show a different language profile to controls, if matched similarly for education and other significant factors.
\end{itemize}

\subsection{Research Challenges}

@@

\section{Project Proposal}

\subsection{Experiment 1 - Pilot Study of Neuropsychological Tests}
\subsubsection{Methods}

@@

\subsubsection{Results}
\subsubsection{Discussion}
\subsubsection{Conclusion}
\subsection{Experiment 2 - Initial Study of the Presidents corpus}
\subsubsection{Background}
\subsubsection{Methods}

@@

\subsubsection{Results}
\subsubsection{Discussion}
\subsubsection{Conclusion}
\subsection{Chapter 3: Experiments with existing datasets: Presidential Press Conferences, Three Authors - Murdoch, Christie and James, DementiaBank dataset.}
Aim: To replicate and extend the work by Berisha and Liss. To use traditional machine learning techniques to test whether they can differentiate between Reagan, Bush and Trump using similar features. To test to see whether Deep Learning can also differentiate between the three. \newline
Dataset: the presidential press conferences given by Presidents Ronald Reagan, George H.W. Bush and Donald Trump. \newline
Paper: This experiment replicates work done by Berisha and Liss and extends this by adding Donald Trump as an alternative comparison to Ronald Reagan. This experiment will look at the features originally recommended by Berisha and Liss, as well as any others that have potential as discussed in the literature review above. \newline
Aim: To replicate the work by Le et al. To use traditional machine learning techniques to see if it can differentiate between the three authors using similar features. To test to see whether deep learning can also differentiate between the three. To add 'pre-trained' layers from Presidential Debates experiment to see if this improves accuracy. \newline
Dataset is the literary novels of Iris Murdoch, Agatha Christie and P.D. James. \newline
Paper:
Aim: To replicate and extend the work of Orimaye et al. To use traditional machine learning techniques to see if it can categorise people into dementia or healthy categories using the samples given. To see whether can use pre-trained layer from presidential debates to improve accuracy. It doesn't seem relevant to use pretrained layer from Experiment four, but it might be worth looking at the results of this. \newline
Dataset: The DementiaBank Corpus. \newline
Paper: To replicate and extend the work of Orimaye et al. To compare traditional machine learning techniques (already completed by Orimaye) and deep learning. \newline
\subsection{Chapter 4: Validation of new repeatable neuropsychological battery for detecting MCI.}
Aim: \newline
Dataset: To be created.

\subsection{Chapter 5: Longditunal Study}
\subsubsection{Dataset creation}
The aim is to recruit a set of 50 participants with MCI / Early Dementia (diagnosed or undiagnosed) and a set of 50 controls with a similar age and broadly similar educational background (a known potential confound in assessing cognitive decline). The rationale for recruiting this number is that there will be an expected drop out rate and so
Paper:
\subsubsection{Ethics}
Need to discuss confidentiality right at the start of each conversation and to reiterate at every meeting. I will also provide an information sheet which details the confidentiality statement. The spoken confidentiality statement will be as follows.
"Anything you say to me today will be confidential unless I feel that there is a risk to you, or someone else. If this is the case, I may need to speak to or write to third parties such as your GP. Is this ok?" \newline
\par
\subsubsection{What if someone reveals suicidal ideation or other risks?}
As part of the battery of tests for this experiment, there is an opportunity for a participant to reveal that he or she has had suicidal ideation in the past two weeks (Patient Health Questionnaire, Question 9). This is something to explore further. As I have been trained in psychological risk assessments, it seems appropriate to carry out a further risk assessment.  Optionally with the tape recorder turned on, or off as requested.  Any text recorded during the risk assessment will not be included in the language analysis. Following this, some information about local services such as Improving Access to Psychological Therapies (IAPT), Samaritans and other appropriate services will be provided. A discussion will then be made about whether it is appropriate to notify the GP of the disclosure. In the event that there is imminent risk to self or others, a discussion will be made about who I need to contact. (Crisis Team and GP). In the event that the participant scores 0 on Q9 of the PHQ, no risk assessment will be undertaken. \newline

\section{Future Work}
\subsection{Introduction}

\subsection{Project Plan}

\subsection{Risk Assessment}

\subsection{Ethical Considerations}

\subsection{Thesis - Provisional Table of Contents}
In this section I provide a provisional table of contents based on the work I have either completed or proposed above.
\begin{enumerate}
	\item Introduction
	\begin{enumerate}
		\item Overview
		\item Background and Context
		\item Aims and Objectives
		\item Challenges and Achievements
	\end{enumerate}
	\item Literature Review
	\begin{enumerate}
		\item Introduction
		\item Assessing Mild Cognitive Impairment using cognitive tests.
		\item Assessing Mild Cognitive Impairment using Language.
		\item Natural Language Processing and Machine Learning
		\item Conclusion
	\end{enumerate}
	\item Further Exploration of the Presidents Corpus
	\begin{enumerate}
		\item Introduction
		\item Methodology
		\item Results
		\item Discussion
		\item Conclusion
	\end{enumerate}
	\item Validation of a Repeatable Neuropsychological Battery of Tests
	\begin{enumerate}
		\item Introduction
		\item Methodology
		\item Pilot Study Results
		\item Longitundinal Study results
		\item Discussion
		\item Conclusion
	\end{enumerate}
	\item Longitudnal Study
	\begin{enumerate}
		\item Introduction
		\item Methodology
		\item Experiment 1 Results
		\item Experiment 2 Results
		\item Discussion
		\item Conclusion
	\end{enumerate}
	\item General Discussion
	\begin{enumerate}
		\item Discussion
		\item Future Work
		\item Conclusion
	\end{enumerate}
\end{enumerate}

\section{Conclusion}
